---
title: |
    | Problem Set 4
    | Statistical Tests
    | POLS3316: Statistics for Political Scientists
    | Due November 15, 2023 at 11:59pm
author: Yoselin Monterroso
format: 
        html:
                self-contained: true
---


## Use this work file if working in R

It is recommended to use this .qmd file to complete your work if working in R. Simply edit your name in the second line. 

## General instructions

`#` indicates that the line is a comment and does not need to be edited.

The significance level for all tests is .05 unless otherwise specified.

Type only code in the code chunks. Do not type answers as comments in the code chunks.

Type answers in the text chunks.

You can get help with the functions using the ? command in the Console (lower left)

To complete Bonus Problem 7, you will need to install a package and load a library. Instructions are given. You will not be able to get help for the z.test using the ? until the package is installed.

## Problem 1: Chi-square

Complete a Chi-square test of independence on the following data using the R function `chisq.test()`.

| | 1 | 2 | 
| --- | --- | --- |
| A | 10 | 20 |
| B | 15 | 25 |

```{r}
# Your code here
A <- c(10,20)

B <- c(15,25)

matrix1 <- cbind(A,B)
chisq.test(matrix1 )


```
X-squared:0.011667

p-value:0.914

Is the test significant? Why or why not?
Yes, because the degree of freedom is 1, which means that the p-value (0.914) is below
the significance level 0.05 (3.841). 

## Problem 2: Chi-square

Complete a Chi-square test of independence on the following data using the R function `chisq.test()`.

| | 1 | 2 | 3 | 4 | 
| --- | --- | --- | --- | --- |
| A | 10 | 20 | 30 | 40 |
| B | 15 | 25 | 35 | 45 |
| C | 20 | 30 | 40 | 50 |

```{r}
# Your code here
A2<- c(10,20,30,40)

B2 <- c(15,25,35,45)

C2 <- c(20,30,40,50)

matrix2 <- cbind(A2,B2,C2)
chisq.test(matrix2)


```

Chi-squared:1.2408

p-value:0.9748

Is the test significant? Why or why not? 
Yes, because the degree of freedom is 6, which means that the p-value (1.2408) is below
the significance level 0.05 (12.592).

## Problem 3: t-test

Complete a paired sample t-test on the following data using the R function `t.test()`. Complete a two-tailed test.

| | Before | After |            
| --- | --- | --- |
| A | 10 | 20 |
| B | 24 | 25 |
| C | 18 | 30 |
| D | 25 | 35 |
| E | 41 | 40 |
| F | 35 | 45 |



```{r}

# Your code here
A3<- c(10,20)
B3 <- c(24,25)
C3 <- c(18,30)
D3<- c(25,35)
E3 <- c(41,40)
F3 <- c(35,45)

matrix3 <- cbind(A3,B3,C3,D3,E3,F3)
t.test(matrix3)


```



t-score:9.5626

p-value:1.154

Is the test significant? Why or why not?
No, it is not significant because it falls above the p-value.

## Problem 5: t-test

Complete a one-sample t-test on the following data using the R function `t.test()`. Complete a two-tailed test.

| | 1 | 2 | 3 | 4 | 5 | 6 | 7 | 8 | 9 | 10 |
| --- | --- | --- | --- | --- | --- | --- | --- | --- | --- | --- |
| A | 10 | 20 | 30 | 40 | 50 | 60 | 70 | 80 | 90 | 100 |

```{r}
# your code here
A4<- c(10,20,30,40,50,60,70,80,90,100)

matrix4 <- cbind(A4)
t.test(matrix4)

```

t-score:5.7446

p-value:0.0002782

Is the test significant? Why or why not?
No, because it falls above the p-value.

## Problem 6: Bonus

Complete a one-way ANOVA on the following data using the R function `aov()`. Complete a two-tailed test.

| | 1 | 2 | 3 | 4 |
| --- | --- | --- | --- | --- |
| A | 10 | 20 | 30 | 40 |
| B | 15 | 25 | 35 | 45 |
| C | 20 | 30 | 40 | 50 |

```{r}
# your code here
A5<- c(10,20,30,40)
B5 <- c(15,25,35,45)
C5 <- c(20,30,40,50)

matrix5 = cbind(A5,B5,C5)
combined_matrix5 = data.frame(matrix5)
stacked_matrix5 = stack(combined_matrix5)
Aov_matrix5 = aov(values ~ ind, data = stacked_matrix5)
summary(Aov_matrix5)
```

Is the test significant? Why or why not?
Yes, because it falls below the p-value. 

## Problem 7: Bonus

Complete a z-test on the following data using the R function `z.test()`. Complete a two-tailed test.

| | obs 1 | obs 2 | obs 3 | obs 4 | obs 5 | obs 6                
| --- | --- | --- | --- | --- | --- | --- |
| A | 10 | 20 | 30 | 40 | 50 | 60 |
| B | 15 | 25 | 35 | 45 | 55 | 65 |              


```{r}
# To run a z.test, you will need to install the BSDA package and load the library

# To do this, remove the # from the following two lines of code and run them

#install.packages("BSDA")
#library(BSDA)


# Your code here
A6 = c(10,20,30,40,50,60)
B6 = c(15,25,35,45,55,65)

#couldn't figure out how to do this one. 
```
z-score here:

p-value here:

Is the test significant? Why or why not?





#Had help figuring out how to do the last two questions throught video(s)
# https://www.youtube.com/watch?v=fT2No3Io72g
